---
title: "HrtDisease tutorial"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Install needed packages
## Load packages
```{r}
library(tidyverse)	
library(readr)
library(e1071)
library(corrplot)
library(ggcorrplot)
library(MASS)
library(Hmisc)
library(DataExplorer)
library(GGally)
```
## Import the dataset, heartdisease.csv
```{r}

```

## Find the structure and data types of the variables.
```{r}

```

## Missing values
Plot the missing values using the DataExplorer::plot_missing() function.
```{r}

```

## Transform variables
Create a data object,  hd2. 
a. Use mutate() to create new variables, “sex”, “fbs”, “exang”, “cp”, and “restecg” and using the if_else function.
b. Create factors of “slope”, “ca”, “thal” 
(Fill in the missing information to make the script work)
```{r}
hd2 <- %>% 
  mutate(sex = if_else(" "),
         fbs = if_else(fbs == 1, ">120", "<=120"),
         exang = if_else(   ),
         cp =  if_else(cp == 1, "ATYPICAL ANGINA",
               if_else(cp == 2, "NON-ANGINAL PAIN", "ASYMPTOMATIC")),
         restecg = if_else(restecg == 0, "NORMAL",
               if_else(restecg == 1, "ABNORMALITY", "PROBABLE OR DEFINITE")),
         slope = as.factor(slope)
         ca = 
         thal = 
         target = if_else(target == 1, "YES", "NO")
         ) %>% 
  mutate_if(is.character, as.factor) %>% 
  dplyr::select(target, sex, fbs, exang, cp, restecg, slope, ca, thal, everything())
```

## Summary statistics
## Find the summary statistics of the dataset, hd2.
```{r}

```

## Using the dataframe, hd2, create boxplots of columns “age”:”oldpeak” using base graphics.

```{r}

```

## Data Visualization
# Create a barplot for the target variable, Heart Disease. Add an x-axis label, “Heart Disease” and y-axis label, “Count”.
```{r}
ggplot(data2, aes(, fill=data2$target)) + 
  geom_bar() +
  +
  +
  ggtitle("Analysis of Presence and Absence of Heart Disease") +
  scale_fill_discrete(name = "Heart Disease", labels = c("Absence", "Presence"))
```

## Find the proportion (frequency) of the dataframe, hd2, where target equals “NO” and target equals “YES”. Use the $ syntax.
```{r}

```

## Group by age, count, n, the frequency of the “age” variable. Filter values where count > 10. Add an x-axis label, “age” and y-axis, “AgeCount”. 
```{r}
data2 %>% 
  %>% 
  %>% 
  %>% 
  ggplot()+
  geom_col(aes(), fill = "navyblue") +
  ggtitle("Age Analysis") +
            +
            
```

## Create a bar graph of the variable, “cp” representing levels of the target variable. Fill in the blanks.
```{r}
ggplot(data2, aes(  , fill =   )) +
  geom_bar(position = "fill") +
  ggtitle("cp")
```

## Let’s look at the distribution of Male and Female population across the “age” parameter. Use the histogram geometry. 
```{r}
data2 %>%
  ggplot(aes( , )) +
      () +
  xlab("Age") + 
  ylab("Number") +
  guides(fill = guide_legend(title = "Gender"))
```

## Let’s look at how cholesterol levels are represented in the data. Using the input dataframe, hd2, create a scatterplot of “age” and “chol”. Use options color to color the points by sex and size to represent the size of the point by the cholesterol level.
```{r}
     
    %>%
  ggplot(aes(x=,y=,color=  , size=  )) +
  geom_  (alpha=0.7) +
  xlab( ) +
  ylab( ) +
  guides(fill = guide_legend(title = "Gender"))
```

## Let’s compare blood pressure across pain type. Using the input dataframe, hd2, create multiple boxplots of “trestbps” by “sex”. Lastly use facet_grid() with the “cp” variable.
```{r}
    %>%
  ggplot(aes(x=,y=  )) +
  geom_ (fill="darkorange") +
  xlab( ) +
  ylab( ) +
  facet_grid(~cp)
```

## Let’s compare cholestrol across pain type. Create multiple boxplots of “chol” by “sex”. Fill the box using a hex code. Lastly use facet_grid() with the “cp” variable.
```{r}
data2 %>%
  ggplot(aes(x=,y=  )) +
  geom_  (fill="#D55E00") +
  xlab(" ") +
  ylab(" ") +
  facet_grid(~cp)
```

## Find the correlations of columns 10:14. Map the correlations to an object “corr_heart”. Use the dataframe, corr_heart, to create a corrplot, keeping only the upper diagonal. Use method=ellipse. 
```{r}
cor_heart <- cor(  )
cor_heart

#run next script as is
corrplot(cor_heart, method = "ellipse", type="upper",) 

## Option #2: (run as is) 
ggcorrplot(cor_heart,lab = T)
ggcorr(cor_heart, label = T, label_round = 2)
```

## Use the library, gridExtra, to create multiple bar plots of “sex”, “fbs”, “exang”. Use the data frame, hd2, as the input data frame. (Run script as is)
```{r}
library(gridExtra)

grid.arrange(
  ggplot(data2, aes(x = sex, fill = target)) +
  geom_bar(position = "fill"),
  
  ggplot(data2, aes(x = fbs, fill = target)) +
  geom_bar(position = "fill"),
  
  ggplot(data2, aes(x = exang, fill = target)) +
  geom_bar(position = "fill"), nrow = 3  )
```

## Create scatterplots using “age” and “trestbps”. Use the options color() and shape() to color code and size the points of the “target” variable. Include a best fit line.
```{r}

```
## Create scatterplots using “age” and “chol”. Use the options color() and shape() to color code and size the points of the “target” variable. Include a best fit line.
```{r}

```
## Create scatterplots using “age” and “thalach”. Use the options color() and shape() to color code and size the points of the “target” variable. Include a best fit line.
```{r}

```
## Create scatterplots using “age” and “oldpeak”. Use the options color() and shape() to color code and size the points of the “target” variable. Include a best fit line.
```{r}

```
